Clustering time series based on probability distributions across temporal granularities

### 2021-11-18 main.Rmd
fixing plot reference

Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:parcoord-application}. The t-SNE visualization suggests that there are four well-separated clusters. It is possible that because the representation is only 2D, that the fifth group from the original prototype selection is distinctly different in high dimensions. The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.  JS-RS would appear to match the original prototypes with the five cluster solution, but it actually differs. Even though the $sindex$ for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes. The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.  <!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.-->

@@

<!-- In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances. The scales are very different from those used in JS-based clustering. The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$. Thus, there are few choices available for the number of clusters.--> Figure \ref{fig:parcoord-application} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution. XXX REWORK THIS WHEN NEW CLUSTER RESULTS ARE SHOWN<!--The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities. The variables are sorted according to their separation across classes (rather than their overall variation between classes). This means that $moy$ is the most important variable in distinguishing the groups, followed by $hod$ and $wnwd$. The two customers (id: 16 and 18) that have significant differences between $wnwd$ categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers. Group-4 has a higher $wpd$ for $hod$ than $moy$ or $wnwd$. Group-2 and Group-3 have the most distinct patterns across $moy$. Group-1 is a mixed group that has strong patterns on at least one of the three variables. When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to $moy$, $hod$ and $wnwd$ are the same. The findings vary from JS-based clustering, yet it is a helpful grouping.-->

@@

```{r parcoord-application, fig.cap="Summary of the four cluster results from WPD, shown as a parallel coordinate plot and table. XXX TO BE REVISED WHEN DIFFERENT CLUSTER SOLUTION IS SHOWN", message=FALSE, warning=FALSE, fig.height=6}
#include_graphics("figs/parcoord-application.png")
```

>>>>

Clustering of the $24$ prototypes <!--and the data subset consisting of $353$ customers--> was conducted with all three distances, JS-NQT, JS-RS and WPD, and is summarised in Figures \ref{fig:opt-cluster-tsne-jsd}, \ref{fig:groups-4and5} and \ref{fig:summary-plot-wpd}. The t-SNE visualization suggests that there are four well-separated clusters. It is possible that because the representation is only 2D, that the fifth group from the original prototype selection is distinctly different in high dimensions. The $sindex$ plots for the three methods indicate some disagreement: JS-NQT suggests $3$, JS-RS suggests $2$ or $5$ and WPD suggests $3$ or $5$.  JS-RS would appear to match the original prototypes with the five cluster solution, but it actually differs. Even though the $sindex$ for JS-NQT suggests three clusters, the five cluster solution more closely matches the original prototypes. The WPD clustering provides a different grouping of the customers, and even though it disagrees with the original prototypes it is a useful grouping.  <!--While plotting the clusters, RS-based scaling is applied to each customer, so that no customer dominates the shape of the summarized clusters because of magnitude. The plotting scales are not displayed since we want to emphasize comparable patterns rather than scales. The idea is that a customer in a cluster may have low total energy usage, but their behavior may be quite similar to a customer with high usage with respect to distributional pattern or significance across cyclic granularities.-->

@@

<!-- In Figure \ref{fig:opt-cluster-tsne-jsd}, the $sindex$ is plotted against the number of clusters for clustering based on wpd-based distances. The scales are very different from those used in JS-based clustering. The $sindex$ falls from around $20$ when $k=3$ to $19$ when $k=4$ and further decreases to $17.5$ with $k>7$. Thus, there are few choices available for the number of clusters.--> Figure \ref{fig:summary-plot-wpd} whos the wpd values for the three different granularities of the 24 customers, coloured by the 6 cluster solution. XXX REWORK THIS WHEN NEW CLUSTER RESULTS ARE SHOWN<!--The groupings with $k=4$ is shown through a parallel coordinate plot with the three significant cyclic granularities. The variables are sorted according to their separation across classes (rather than their overall variation between classes). This means that $moy$ is the most important variable in distinguishing the groups, followed by $hod$ and $wnwd$. The two customers (id: 16 and 18) that have significant differences between $wnwd$ categories (in Figure \ref{fig:hod-ind-group-png}) stand out from the rest of the customers. Group-4 has a higher $wpd$ for $hod$ than $moy$ or $wnwd$. Group-2 and Group-3 have the most distinct patterns across $moy$. Group-1 is a mixed group that has strong patterns on at least one of the three variables. When $k=3$ is used, Group-2 and Group-3 are merged because their patterns relative to $moy$, $hod$ and $wnwd$ are the same. The findings vary from JS-based clustering, yet it is a helpful grouping.-->

@@

<!-- # ```{r summary-plot-wpd, fig.cap="Summary of the four cluster results from WPD, shown as a parallel coordinate plot and table. XXX TO BE REVISED WHEN DIFFERENT CLUSTER SOLUTION IS SHOWN", message=FALSE, warning=FALSE, fig.height=6} -->
<!-- # #include_graphics("figs/summary-plot-wpd.png") -->
<!-- # ``` -->
